Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-07-10 paper_jasa.tex
removed appendix, and some blue color comments

Each subject viewed at least 10 lineups and we allowed them to provide more evaluations if they wanted. Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed (actual numbers are given in the Appendix). For each difficulty level a specific number of lineups were randomly picked to show for evaluation. This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$. To satisfy this plan we needed to recruit at least 300 subjects.

@@

\begin{table}[hbtp]
\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets. Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information. \blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
\begin{center}
\begin{tabular}{ccrrcc}
   \hline 
 Experiment & Subject & Male & Female & Responses &Lineup\\ 
    \hline
1 & 239 & 121 & 107 & 2249 &  60 \\ 
  2 & 351 & 185 & 164 & 3636 &  70 \\ 
  3 & 155 & 103 &  52 & 1511 &  29 \\  
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

@@

%\end{multicols}
\section*{Appendix}

@@

\begin{figure}[hbtp]
   \centering
       \scalebox{0.70}{\includegraphics{power_screening.pdf}}
       \caption{The overall power estimated from model \ref{eqn:mixed} are shown for various data sets obtained by applying 6 screening criteria.  The corresponding power curve for conventional test is shown in black solid lines for comparison.}
       \label{fig:power_screening}
\end{figure}

\begin{figure}[hbtp]
   \centering
       \scalebox{0.60}{\includegraphics{power_screening_subject.pdf}}
      \caption{The subject-specific power estimated from model \ref{eqn:mixed} are shown for various data sets obtained by applying 6 screening criteria.  The corresponding power curve for conventional test is shown in pink lines for comparison.}
       \label{fig:power_screening_subject}	
\end{figure}

\begin {abstract} 
Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently, there were no formal visual methods in place for determining statistical significance of findings. This changed when Buja et al. (2009) conceptually introduced the lineup protocol for formal tests of visual findings. In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
\end {abstract}

>>>>

Each subject viewed at least 10 lineups and we allowed them to provide more evaluations if they wanted. Depending on the parameter combinations we group the lineups in different difficulty levels such as easy, medium, hard and mixed (actual numbers are given in the supplementary material). For each difficulty level a specific number of lineups were randomly picked to show for evaluation. This number is chosen so that total number of evaluations for each lineup for that group exceed the threshold $n_{\gamma}$. To satisfy this plan we needed to recruit at least 300 subjects.

@@

\begin{table}[hbtp]
\caption{Number of subjects, gender and total lineups seen, and number of distinct lineups for all three experimental data sets. Note that in some of the lines the number of male and female participants does not add up to the total number of participants due to missing demographic information. }%\blue{Notice that we do not have data on one lineup (so we have only 29 lineups) for experiment 3 due to the screening process.}}
\begin{center}
\begin{tabular}{ccrrcc}
   \hline 
 Experiment & Subject & Male & Female & Responses &Lineup\\ 
    \hline
1 & 239 & 121 & 107 & 2249 &  60 \\ 
  2 & 351 & 185 & 164 & 3636 &  70 \\ 
  3 & 155 & 103 &  52 & 1511 &  29 \\  
   \hline
\end{tabular}
\end{center}
\label{tbl:summary}
\end{table}

@@

%\end{multicols}

\section*{Appendix}

@@

%\begin {abstract} 
%Statistical graphics play a crucial role in exploratory data analysis, model checking and diagnosis. Until recently, there were no formal visual methods in place for determining statistical significance of findings. This changed when Buja et al. (2009) conceptually introduced the lineup protocol for formal tests of visual findings. In this paper this is taken a step further by refining the terminology of visual inference, framing the protocol in a context that allows direct comparison with conventional tests, in scenarios when a conventional test exists, and it is used to compare the performance of the lineup protocol against conventional statistical testing in the scenario of fitting linear models. A human subjects experiment is conducted using simulated data to provide controlled conditions. Results suggest that the lineup protocol provides results comparable to the conventional tests, out-performs them when data is contaminated, and, there may be some super-individuals who yield better power than the conventional test even in the most difficult trials.
%\end {abstract}
